Effect of quality-adjusted production function inputs

### 2015-06-05 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Incorporating suggestions from Randy.

Parameter estimation is accomplished with \texttt{formula}s of the form
%
\begin{equation} \label{eq:formulas}
  y \sim x_1 + x_2 + x_3 + time,
\end{equation}
%
where $y$ is economic output and 
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of
capital ($k$), labor ($l$), and energy ($e$).
(The factors of production may be quality-adjusted.)
$time$ is the time variable, usually \texttt{iYear}. 
The \texttt{formula}s are used in a fitting function such as 
\texttt{cdModel} to create a \texttt{model} object.
The function \texttt{naturalCoef} extracts reasonable
coefficients from \texttt{model} objects.

Note that in the fitting process for the Cobb-Douglas equation,
$\texttt{scale} = \theta$, 
$\alpha_1 = \alpha$,
$\alpha_2 = \beta$, and
$\alpha_3 = \gamma$,
when the \texttt{formula} is specified as 
%
\begin{equation} \label{eq:formula_kle}
  \texttt{y} \sim \texttt{k + l + e + time}.
\end{equation}
%
Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity.

>>>>

% Parameter estimation is accomplished with \texttt{formula}s of the form
% %
% \begin{equation} \label{eq:formulas}
%   y \sim x_1 + x_2 + x_3 + time,
% \end{equation}
% %
% where $y$ is economic output and 
% $x_1$, $x_2$, and $x_3$ are factors of production
% and permutations of
% capital ($k$), labor ($l$), and energy ($e$).
% (The factors of production may be quality-adjusted.)
% $time$ is the time variable, usually \texttt{iYear}. 
% The \texttt{formula}s are used in a fitting function such as 
% \texttt{cdModel} to create a \texttt{model} object.
% The function \texttt{naturalCoef} extracts reasonable
% coefficients from \texttt{model} objects.
% 
% Note that in the fitting process for the Cobb-Douglas equation,
% $\texttt{scale} = \theta$, 
% $\alpha_1 = \alpha$,
% $\alpha_2 = \beta$, and
% $\alpha_3 = \gamma$,
% when the \texttt{formula} is specified as 
% %
% \begin{equation} \label{eq:formula_kle}
%   \texttt{y} \sim \texttt{k + l + e + time}.
% \end{equation}
% %
% Note that for the CES model, $\texttt{scale} = \gamma$ and is expected to be close to unity.

The models being fit can be described by the algebraic form of the model 
(Cobb-Douglas, CES, etc.) and a formula that enumerates 
which data variables are being used in which roles to fit the model.  
A formula of the form
%
\begin{equation} \label{eq:formulas-3-factor}
y \sim x_1 + x_2 + x_3 + t,
\end{equation}
%
or 
%
\begin{equation} \label{eq:formulas-2-factor}
y \sim x_1 + x_2 + t,
\end{equation}
%
describes the economic output variable ($y$, usually \texttt{iGDP},
                                        indexed GDP) 
and the factors of production ($x_1$, $x_2$, and $x_3$,
                               which will be some measure of capital, labor, and energy, 
                               but perhaps not in that order), 
and a time variable ($t$, usually \texttt{iYear}, 
                     the number of years since the beginning of data collection).  
All of the models assume an error term that is additive on the 
logarithmic scale and are fit by the method of least squares.  
Model fitting provides estimates for all parameters in the model.
